How to compute the fusion product of MV cycles in type A

### 2021-04-10 draft.tex

\begin{example}
Let $G = \GL_3$ and consider the tableaux
$$\tau_1 = \young(12,3) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2=\young(13,2).$$
The MV cycles corresponding to $\tau_1$ and $\tau_2$ are both isomorphic to $\PP^2$.
The weights are $\lambda_1 = \lambda_2 = (2,1,0)$ and $\mu_1 = \mu_2 = (1,1,1)$. Hence the matrix in consideration is 
\[
X = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc}
    0 & 1 & & & & \\
     & s & a_1 & a_2 & b_1 & b_2 \\
     & & 0 & 1 & & \\
     & & & s & b_3 & b_4 \\
     & & & & 0 & 1 \\
     & & & & & s
\end{BMAT}
\right].
\]

@@

Hence we consider the quotient
$$\frac{\CC[a_1,a_2,b_1,b_2,b_3,b_4,s]}{\langle a_1+sa_2,b_3,s \rangle} \cong
\frac{\CC[a_1,a_2,b_1,b_2,b_3,b_4]}{\langle a_1,b_3 \rangle}.$$
We see already that the $0$-fibre is irreducible. The tableau corresponding to the ideal $\langle a_1, b_3 \rangle$ is 
$$\young(1123,23)$$
which corresponds to the MV cycle that is the product of the MV cycles corresponding to $\tau_1$ and $\tau_2$.
\end{example}

\begin{example}
Let $G = SL_3$ and consider the tableaux
$$\tau_1 = \young(12) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2 = \young(11,23).$$
The MV cycles corresponding to the above tableaux are each isomorphic to $\PP^1$.
We have the weights to be $\lambda_1 = (2,0,0)$, $\lambda_2 = (2,2,0)$, $\mu_1 = (1,1,0)$, and $\mu_2 = (2,1,1)$. The matrix under consideration is therefore
\[
X = \left[\begin{BMAT}(e){ccc;cc;c}{ccc;cc;c}
    0 & 1 & & & & \\
     & 0 & 1 & & & \\
     & -s^2 & 2s & a_1 & a_2 & b_1 \\
     & & & 0 & 1 & \\
     & & & & s & b_2 \\
     & & & & & s
\end{BMAT}
\right].
\]
There are no relations from looking at the submatrix $X_1$. From the submatrix $X_2$, we require $$a_1 + sa_2 = 0.$$
From $X$, we further obtain the relations $$a_2b_2 + sb_1 = 0, a_1b_2 + s^2b_1 = 0.$$
Then the quotient ring we find is
$$\frac{\CC[a_1,a_2,b_1,b_2,s]}{\langle a_1+sa_2, a_2b_2+sb_1, a_1b_2+s^2b_1, s\rangle} \cong \frac{\CC[a_1,a_2,b_1,b_3,s]}{\langle a_1, a_2\rangle \cap \langle a_1, b_2 \rangle}.$$
We see that the 0-fibre is reducible with two components. The tableau corresponding to each ideal is
$$\begin{array}{cc} \vspace{1mm}
    \young(1113,22) & \text{for } \langle a_1, a_2 \rangle \\ 
    \young(1112,23) & \text{for } \langle a_1, b_2 \rangle.
\end{array}$$
Each of these tableaux corresponds to a MV cycle isomorphic to $\PP^2$ so the generic fibre $\PP^1 \times \PP^1$ degenerates to two copies of $\PP^2$.
\end{example}

\begin{example}
Let $G = SL_3$ and consider the following tableaux:
$$\tau_1 = \young(1122) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2 = \young(1111,2233).$$
The MV cycles corresponding to these tableaux are each isomorphic to $\PP^1 \times \PP^1$. 
The weights are then $\lambda_1 = (4,0,0)$, $\lambda_2 = (4,4,0)$, $\mu_1 = (2,2,0)$, and $\mu_2 = (4,2,2)$.
Then we are considering the following matrix:
\[
X = \left[\begin{BMAT}(e){cccccc;cccc;cc}{cccccc;cccc;cc}
    0 & 1 & & & & & & & & & & \\
     & 0 & 1 & & & & & & & & & \\
     & & 0 & 1 & & & & & & & & \\
     & & & 0 & 1 & & & & & & & \\
     & & & & 0 & 1 & & & & & & \\
     & & -s^4 & 3s^3 & -6s^2 & 4s & a_1 & a_2 & a_3 & a_4 & b_1 & b_2 \\
     & & & & & & 0 & 1 & & & & \\
     & & & & & & & 0 & 1 & & & \\
     & & & & & & & & 0 & 1 & & \\
     & & & & & & & & -s^2 & 2s & b_3 & b_4 \\
     & & & & & & & & & & 0 & 1 \\
     & & & & & & & & & & -s^2 & 2s
\end{BMAT}
\right].
\]
By computer, it can be checked that we obtain the following 3 ideals:
$$
\begin{array}{cl}
    I_1 = & \langle a_1, a_2, a_3, a_4, s \rangle \\
    I_2 = & \langle a_1, a_2, b_3, b_4, s \rangle \\
    I_3 = & \langle a_1, a_2, b_3^2, a_4b_3 + a_3b_4, a_3b_3, a_3^2, s \rangle.
\end{array}
$$
It can be checked that $I_3$ has multiplicity 2 \rcom{Not sure how to make this precise and how to check it.} so the 0-fibre has 4 components, two of which are the same. The tableau corresponding to each ideal is
$$
\begin{array}{cc}\vspace{1mm}
    \young(11111133,2222) & \text{for } I_1 \\ \vspace{1mm}
    \young(11111122,2233) & \text{for } I_2 \\ \vspace{1mm}
    \young(11111123,2223) & \text{for } I_3.
\end{array}
$$
The MV cycle for each ideal is isomorphic to $\PP^2 \times \PP^2$.
\end{example}

\begin{example}
Let us redo example 3 but with non-dominant weights $\mu_i$ such that $\mu_1 + \mu_2$ is still dominant. Then consider the tableaux
$$\tau_1 = \young(22) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2 = \young(11,33).$$
Then the weights are $\lambda_1 = (2,0,0)$, $\lambda_2 = (2,2,0)$, $\mu_1 = (0,2,0)$, and $\mu_2 = (2,0,2)$. Our matrix $X$ is now
\[
X = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc}
    0 & 1 & & & & \\
    -s^2 & 2s & a_1 & a_2 & b_1 & b_2 \\
     & & 0 & 1 & & \\
     & & & 0 & b_3 & b_4 \\
     & & & & 0 & 1 \\
     & & & & -s^2 & 2s
\end{BMAT}
\right].
\]
Using a computer, the ideals we now get are
$$
\begin{array}{cl}
    I_1 = & \langle a_1, a_2 \rangle \\
    I_2 = & \langle b_3, b_4 \rangle \\
    I_3 = & \langle b_3^2, a_2b_3 + a_1b_4, a_1b_3, a_1^2 \rangle. 
\end{array}
$$
The tableaux corresponding to each ideal is
$$
\begin{array}{cc}\vspace{1mm}
    \young(1133,22) & \text{for } I_1 \\ \vspace{1mm}
    \young(1122,33) & \text{for } I_2 \\ \vspace{1mm}
    \young(1123,23) & \text{for } I_3
\end{array}
$$
and the MV cycles corresponding to each are the same as before.
\[
{\lambda'}^{(i)}    
\]  
\end{example}

>>>>

\begin{example}
Let $G = \GL_3$ and consider the tableaux
$$\tau_1 = \young(12,3) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2=\young(13,2).$$

Lusztig data: $(010)$ for $\tau_1$, $(101)$ for $\tau_2$

Dimensions: $(1,0,-1)$ for both.

The MV cycles corresponding to $\tau_1$ and $\tau_2$ are both isomorphic to $\PP^2$.
The weights are $\lambda_1 = \lambda_2 = (2,1,0)$ and $\mu_1 = \mu_2 = (1,1,1)$. Hence the matrix in consideration is 
\[
X = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc}
    0 & 1 & & & & \\
     & s & a_1 & a_2 & b_1 & b_2 \\
     & & 0 & 1 & & \\
     & & & s & b_3 & b_4 \\
     & & & & 0 & 1 \\
     & & & & & s
\end{BMAT}
\right].
\]

@@

Hence we consider the quotient
$$\frac{\CC[a_1,a_2,b_1,b_2,b_3,b_4,s]}{\langle a_1+sa_2,b_3,s \rangle} \cong
\frac{\CC[a_1,a_2,b_1,b_2,b_3,b_4]}{\langle a_1,b_3 \rangle}.$$
We see already that the $0$-fibre is irreducible. The tableau corresponding to the ideal $\langle a_1, b_3 \rangle$ is 
$$\young(1123,23)$$
which corresponds to the MV cycle that is the product of the MV cycles corresponding to $\tau_1$ and $\tau_2$.

Lusztig datum: $(111)$

Dimension: $(2,0,-2)$
\end{example}

\begin{example}
Let $G = SL_3$ and consider the tableaux
$$\tau_1 = \young(12) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2 = \young(11,23).$$

Lustig data: $(100)$ for $\tau_1$, $(001)$ for $\tau_2$.

Dimensions: $(1,-1,0)$ for $\tau_1$, $(0,1,-1)$ for $\tau_2$.

The MV cycles corresponding to the above tableaux are each isomorphic to $\PP^1$.
We have the weights to be $\lambda_1 = (2,0,0)$, $\lambda_2 = (2,2,0)$, $\mu_1 = (1,1,0)$, and $\mu_2 = (2,1,1)$. The matrix under consideration is therefore
\[
X = \left[\begin{BMAT}(e){ccc;cc;c}{ccc;cc;c}
    0 & 1 & & & & \\
     & 0 & 1 & & & \\
     & -s^2 & 2s & a_1 & a_2 & b_1 \\
     & & & 0 & 1 & \\
     & & & & s & b_2 \\
     & & & & & s
\end{BMAT}
\right].
\]
There are no relations from looking at the submatrix $X_1$. From the submatrix $X_2$, we require $$a_1 + sa_2 = 0.$$
From $X$, we further obtain the relations $$a_2b_2 + sb_1 = 0, a_1b_2 + s^2b_1 = 0.$$
Then the quotient ring we find is
$$\frac{\CC[a_1,a_2,b_1,b_2,s]}{\langle a_1+sa_2, a_2b_2+sb_1, a_1b_2+s^2b_1, s\rangle} \cong \frac{\CC[a_1,a_2,b_1,b_3,s]}{\langle a_1, a_2\rangle \cap \langle a_1, b_2 \rangle}.$$
We see that the 0-fibre is reducible with two components. The tableau corresponding to each ideal is
$$\begin{array}{cccc} \vspace{1mm}
    \young(1113,22) & \text{for } \langle a_1, a_2 \rangle & \text{Lusztig: }(101) & \dim = (1,0,-1)\\ 
    \young(1112,23) & \text{for } \langle a_1, b_2 \rangle & \text{Lusztig: }(010) & \dim = (1,0,-1).
\end{array}$$
Each of these tableaux corresponds to a MV cycle isomorphic to $\PP^2$ so the generic fibre $\PP^1 \times \PP^1$ degenerates to two copies of $\PP^2$.
\end{example}

\begin{example}
Let $G = SL_3$ and consider the following tableaux:
$$\tau_1 = \young(1122) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2 = \young(1111,2233).$$

Lusztig data: $(200)$ for $\tau_1$, $(002)$ for $\tau_2$.

Dimensions: $(2,-2,0)$ for $\tau_1$, $(0,2,-2)$ for $\tau_2$.

The MV cycles corresponding to these tableaux are each isomorphic to $\PP^1 \times \PP^1$. 
The weights are then $\lambda_1 = (4,0,0)$, $\lambda_2 = (4,4,0)$, $\mu_1 = (2,2,0)$, and $\mu_2 = (4,2,2)$.
Then we are considering the following matrix:
\[
X = \left[\begin{BMAT}(e){cccccc;cccc;cc}{cccccc;cccc;cc}
    0 & 1 & & & & & & & & & & \\
     & 0 & 1 & & & & & & & & & \\
     & & 0 & 1 & & & & & & & & \\
     & & & 0 & 1 & & & & & & & \\
     & & & & 0 & 1 & & & & & & \\
     & & -s^4 & 3s^3 & -6s^2 & 4s & a_1 & a_2 & a_3 & a_4 & b_1 & b_2 \\
     & & & & & & 0 & 1 & & & & \\
     & & & & & & & 0 & 1 & & & \\
     & & & & & & & & 0 & 1 & & \\
     & & & & & & & & -s^2 & 2s & b_3 & b_4 \\
     & & & & & & & & & & 0 & 1 \\
     & & & & & & & & & & -s^2 & 2s
\end{BMAT}
\right].
\]
By computer, it can be checked that we obtain the following 3 ideals:
$$
\begin{array}{cl}
    I_1 = & \langle a_1, a_2, a_3, a_4, s \rangle \\
    I_2 = & \langle a_1, a_2, b_3, b_4, s \rangle \\
    I_3 = & \langle a_1, a_2, b_3^2, a_4b_3 + a_3b_4, a_3b_3, a_3^2, s \rangle.
\end{array}
$$
It can be checked that $I_3$ has multiplicity 2 \rcom{Not sure how to make this precise and how to check it.} so the 0-fibre has 4 components, two of which are the same. The tableau corresponding to each ideal is
$$
\begin{array}{cccc}\vspace{1mm}
    \young(11111133,2222) & \text{for } I_1 & \text{Lusztig: }(202) & \dim = (2,0,-2) \\ \vspace{1mm}
    \young(11111122,2233) & \text{for } I_2 & \text{Lusztig: }(020) & \dim = (2,0,-2) \\ \vspace{1mm}
    \young(11111123,2223) & \text{for } I_3 & \text{Lusztig: }(111) & \dim = (2,0,-2).
\end{array}
$$
The MV cycle for each ideal is isomorphic to $\PP^2 \times \PP^2$.
\end{example}

\begin{example}
Let us redo example 3 but with non-dominant weights $\mu_i$ such that $\mu_1 + \mu_2$ is still dominant. Then consider the tableaux
$$\tau_1 = \young(22) \hspace{.5cm} \text{ and } \hspace{.5cm} \tau_2 = \young(11,33).$$
Then the weights are $\lambda_1 = (2,0,0)$, $\lambda_2 = (2,2,0)$, $\mu_1 = (0,2,0)$, and $\mu_2 = (2,0,2)$. Our matrix $X$ is now
\[
X = \left[\begin{BMAT}(e){cc;cc;cc}{cc;cc;cc}
    0 & 1 & & & & \\
    -s^2 & 2s & a_1 & a_2 & b_1 & b_2 \\
     & & 0 & 1 & & \\
     & & & 0 & b_3 & b_4 \\
     & & & & 0 & 1 \\
     & & & & -s^2 & 2s
\end{BMAT}
\right].
\]
Using a computer, the ideals we now get are
$$
\begin{array}{cl}
    I_1 = & \langle a_1, a_2 \rangle \\
    I_2 = & \langle b_3, b_4 \rangle \\
    I_3 = & \langle b_3^2, a_2b_3 + a_1b_4, a_1b_3, a_1^2 \rangle. 
\end{array}
$$
The tableaux corresponding to each ideal is
$$
\begin{array}{cc}\vspace{1mm}
    \young(1133,22) & \text{for } I_1 \\ \vspace{1mm}
    \young(1122,33) & \text{for } I_2 \\ \vspace{1mm}
    \young(1123,23) & \text{for } I_3
\end{array}
$$
and the MV cycles corresponding to each are the same as before.
\end{example}
